---
title: "collatz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{collatz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install
After following the ["Getting Started"](https://skenvy.github.io/Collatz/R/#getting-started)
instructions (which can also be seen in the [README](https://github.com/Skenvy/Collatz/tree/main/R))
to download and or install the package, it can be imported with;
```{r setup, message = FALSE}
library(collatz)
# Optionally
library(gmp)
```
You can jump right into using the functions in this package with integer inputs, or you can use `bigz` or `bigq` from the `gmp` library.
Each function, amongst possible _other_ parameters, comes with the default parameters `(P=2,a=3,b=1)` as optional inputs.
In each case, `P[=2]` is the modulus to check for divisibility of the input by, `a[=3]` is the factor to multiply the input by,
and `b[=1]` is the value to add to the product of `a[=3]` and the input.
There are two basic commands to start with; the `collatzFunction` and the `reverseFunction`.

## collatzFunction
Returns the output of a single application of a Collatz-esque function.
Without `gmp` or parameterisation, we can try something simple like
```{r}
collatzFunction(5)
collatzFunction(16)
```
If we want change the default parameterisation we can;
```{r}
collatzFunction(4, 5, 2, 3)
```
Or if we only want to change one of them
```{r}
collatzFunction(3, a=-2)
```
All the above work fine, but the function doesn't offer protection against overflowing integers by default.
To venture into the world of arbitrary integer inputs we can use an `as.bigz` from `gmp`. Compare the two;
```{r}
collatzFunction(99999999999999999999)
collatzFunction(as.bigz("99999999999999999999"))
```

## reverseFunction
Calculates the values that would return the input under the Collatz function.
Without `gmp` or parameterisation, we can try something simple like
```{r}
reverseFunction(1)
reverseFunction(2)
reverseFunction(4)
```
If we want change the default parameterisation we can;
```{r}
reverseFunction(3, -3, -2, -5)
```
Or if we only want to change one of them
```{r}
reverseFunction(16, a=5)
```
All the above work fine, but the function doesn't offer protection against overflowing integers by default.
To venture into the world of arbitrary integer inputs we can use an `as.bigz` from `gmp`. Compare the two;
```{r}
reverseFunction(99999999999999999999)
reverseFunction(as.bigz("99999999999999999999"))
```
